Distance Sampling in \proglang{R}

### 2016-01-29 paper.Rmd
Added availability plot
to help explain the histogram of detection distances for point transects

Conventional distance sampling assumes the observer is either stood at a point or moving along a line and will only see everything that occurs exactly at that point or on the line. It is then expected that the further away an object is from the point or line (also known as the sampler or transect) the less likely it is that the observer will see it. By recording the distance to each of the detected objects we can fit a model to describe the probability of detection given distance from the sampler, which we refer to as the ‘detection function’. The detection function can be used to infer how many objects were missed and thereby produce estimates of density and/or abundance. Further information on distance sampling, including the theory, various extensions to deal with more complex situations and recent methodological developments can be found at http://distancesampling.org/dbib.html

@@

## Distance sampling in R?

_(LM I wonder if here we should focus on comparing Distance to other R packages offering distance sampling analyses here? Some of the info below could be incorporated into the paragraphs above (I have done this to some extent.. do we need more on distance sampling theory above?)... not done much with this section in case we want to do something different here.)_

@@

@@

```{r lines, fig.height=3, fig.width=7, echo=FALSE, message=FALSE, fig.cap="Left side plot shows an example of a survey of an area containing a population of 500 objects; blue lines indicate sampler placement, red dots indicate detected individuals and grey dots give the locations of unobserved individuals. The right side of the figure shows a histogram of observed distances.\\label{fig:lines}"}
opar <- par(mfrow=c(1,2))
# line and point transect examples
# top row points/samplers in space
# bottom row histograms

#par(opar2)
par(mar=c(4, 4, 1, 2) + 0.1)
hist(detected_distances, main="", xlab="Observed perpendicular distances", freq = FALSE)

@@

<!--

@@

(LT: Only got this far: will do more later...) -->

>>>>

Conventional distance sampling assumes the observer is either stood at a point or moving along a line and will only see everything that occurs exactly at that point or on the line. It is then expected that the further away an object is from the point or line (also known as the sampler or transect) the less likely it is that the observer will see it. By recording the distance to each of the detected objects we can fit a model to describe the probability of detection given distance from the sampler, which we refer to as the ‘detection function’. The detection function can be used to infer how many objects were missed and thereby produce estimates of density and/or abundance.

@@

## Distance sampling

@@

Conventional distance sampling relies on a number of assumptions. Firstly, like plot sampling, distance sampling assumes that the samplers have been located at random within the study area. By ensuring that the samplers are located independently of the objects in the study region we know that the "availability" of objects with respect to line transect samplers should be constant with increasing perpendicular distance. In contrast,

@@

```{r detectdists, fig.height=5.5, fig.width=8, echo=FALSE, message=FALSE, fig.cap="Top panels show an example detection function (left), the expected avaiability of objects (middle) and the resulting destribution of detections (right) for line transects. Bottom panels show an example detection function (left), the expected avaiability of objects (middle) and the resulting destribution of detections (right) for point transects. \\label{fig:detectiondist}"}

opar <- par(mfrow=c(2,3))
#line and point transect examples

x <- seq(0,500, length = 100)
y <- exp(-((x^2)/(2*(200^2))))

#Line transect example
plot(x, y, xlab = "Perpendicular distance (m)", ylab = "Probability of Detection", type = "l", ylim = c(0,1.1), main = "Probability of Detection", xaxp = c(0,1,5))
line.availability <- rep(1,100)
plot(x, line.availability, xlab = "Perpendicular distance (m)", ylab = "Availability", type = "l", ylim = c(0,1.1), xlim = c(0,500), yaxt = "n", main = "Expected Availability")
plot(x,y, xlab = "Perpendicular distance (m)", ylab= "", type = "l", ylim = c(0,1.1), yaxt = "n", main = "Observed Distribution")

#Point Transect example
plot(x, y, xlab = "Radial distance (m)", ylab = "Probability of Detection", type = "l", ylim = c(0,1.1), main = "Probability of Detection", xaxp = c(0,1,5))
point.availability <- 2*pi*x
plot(x, point.availability, xlab = "Radial distance (m)", ylab = "Availability", type = "l", ylim = c(0,3300), xlim = c(0,500), yaxt = "n", main = "Expected Availability")
point.observed <- y*point.availability
plot(x, point.observed, xlab = "Radial distance (m)", ylab= "", type = "l", ylim = c(0,800), yaxt = "n", main = "Observed Distribution")

@@

@@

(LT: Only got this far: will do more later...)
